---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dominionCVR

**Shiro Kuriwaki** and **Jeff Lewis**

<!-- badges: start -->
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
<!-- badges: end -->

Dominion voting machines, used in many places such as Arizona, Orange County, CA, and San Francisco Couty, CA, export their cast vote records in JSON format. This R package provides a common interface to read such data files into tabular form. 
 
 
## Installation

You can install the package from github as follows. You will need a Github Authorization token provided in the `auth_token` argument.

``` r
remotes::install_github("kuriwaki/dominionCVR")
```


## Package and Dependencies

```{r example}
library(dominionCVR)
```


```{r, message=FALSE}
library(tidyverse)
library(jsonlite)
```

## Usage

Currently there is a simple function to read from a JSON file:

```{r}
extract_cvr(path = "data-raw/json/CvrExport_42.json") |> 
  as_tibble()
```

It can read multiple files at the same time and run in multicore settings

```{r}
library(furrr)
plan("multicore")
extract_cvr(path = c("data-raw/json/CvrExport_42.json",
                     "data-raw/json/CvrExport_24940.json",
                     "data-raw/json/CvrExport_AZ-999.json"
                     )) |> 
  as_tibble()
```




## Inside CVR Exports

Each CVR export is a hierarchical data format. 

```{r}
cvr <- fromJSON("data-raw/json/CvrExport_42.json") 

cvr %>% str(max.level = 3, vec.len = 2)
```


The key parts in this top level output is `Sessions$Orignal$Cards`. 

`Orignal` includes the ballot data as taken. Each session includes cards for the ballots.

Each card includes contests. Here we take the 49th card.

```{r}
length(cvr$Sessions$Original$Cards)

# First card
cvr$Sessions$Original$Cards[[49]] %>% str(max.level = 2)
```


Each contest is identified by a _Contest ID_.    

```{r}
cvr$Sessions$Original$Cards[[49]]$Contests[[1]] %>% 
  as_tibble()
```

Each contest ID in a card has itself a table of Marks. A contests' mark is read into R as a list of row-1 dataframes. 

```{r}
cvr$Sessions$Original$Cards[[49]]$Contests[[1]]$Marks %>% str(list.len = 2)
```

Therefore this list can be stacked into a table. 


```{r}
votes <- cvr$Sessions$Original$Cards[[49]]$Contests[[1]]$Marks %>% 
  bind_rows()

as_tibble(votes)
```


Here we finally get to the real part. `IsVote = TRUE`  means that the mark is a vote for the candidate. We can merge this with the candidate metadata to who these candidates are. 




```{r}
# make contest data
cont_df <- fromJSON("data-raw/json/ContestManifest.json")$List %>% 
  rename_all(~ str_c("Contest", .x)) %>% 
  as_tibble()

# merge with candidate data
cand_df <- cont_df %>% 
  select(ContestDescription, ContestId) %>% 
  left_join(fromJSON("data-raw/json/CandidateManifest.json")$List, by = "ContestId") %>% 
  as_tibble()
```


```{r}
labels <- select(cand_df, 
                 contest = ContestDescription, 
                 candidate = Description, 
                 CandidateId = Id)

votes %>% 
  # add name to candidate
  left_join(labels, by = "CandidateId") %>% 
  relocate(contest, candidate, IsVote) %>% 
  as_tibble()
```



## Inside Manifest Files

Metadata about the contests are stored in Manifest files, e.g. for contest codes, use `contestManifest.json` and for candidate codes use `CandidateManifest.json`. These are single-table JSONs that are easy to read, but we provide the function `read_manifest` that exports a tibble and can extract from a zip file too.


```{r}
 read_manifest("data-raw/json/BallotTypeContestManifest.json")
 read_manifest("data-raw/json/BallotTypeManifest.json")
 read_manifest("data-raw/json/CandidateManifest.json")
 read_manifest("data-raw/json/ContestManifest.json")
 read_manifest("data-raw/json/CountingGroupManifest.json")
 read_manifest("data-raw/json/DistrictManifest.json")
 read_manifest("data-raw/json/DistrictPrecinctPortionManifest.json")
 read_manifest("data-raw/json/DistrictTypeManifest.json")
 read_manifest("data-raw/json/ElectionEventManifest.json")
 read_manifest("data-raw/json/OutstackConditionManifest.json")
 read_manifest("data-raw/json/PartyManifest.json")
 read_manifest("data-raw/json/PrecinctManifest.json")
 read_manifest("data-raw/json/PrecinctPortionManifest.json")
 read_manifest("data-raw/json/TabulatorManifest.json")
```
